---
title: "Reproducible Researh - Course project 1"
output: html_document
---


```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

### Loading the data
```{r reading the data}
data <- read.csv("activity.csv")
```

### What is mean total number of steps taken per day?
```{r}
stepsAday <- with(data, tapply(steps, date, sum, na.rm = T))
hist(stepsAday, main = "Number of steps taken each day", xlab = "Number of steps")
```
<br>The mean and median of the total number of steps taken per day:
```{r}
mean(stepsAday, na.rm = T)
median(stepsAday, na.rm = T)
```


### What is the average daily activity pattern?
```{r}
avgs <- aggregate(x = list(steps = data$steps), by = list(intervals = data$interval), FUN = mean, na.rm = TRUE)
library(ggplot2)
ggplot(data = avgs, aes(intervals, steps)) + geom_line() 
```
<br>Which 5-minute interval contains the maximum number of steps?
```{r}
maxInterval <- which.max(avgs$steps)
avgs[maxInterval,]$intervals
```

### Imputing missing values  
The total number of missing values in the dataset:
```{r}
sum(!complete.cases(data))
```
<br> Now let's replace NA values with average steps in that day
```{r}
library(dplyr)
newData <- data

stepsAday <- data.frame(stepsAday)    # average steps per day
stepsAday["date"] <- as.factor(row.names(stepsAday))
names(stepsAday)  <- c("dayAvg", "date") 
rownames(stepsAday) <- c()
if (nrow(stepsAday[is.nan(stepsAday$dayAvg),])>0) stepsAday[is.nan(stepsAday$dayAvg),]$dayAvg <- c(0)

newData <- left_join(newData, stepsAday, by = "date")
names(newData) <- c("steps", "date", "interval", "dayAvg")
newData[is.na(newData$steps),]$steps <- newData[is.na(newData$steps),]$dayAvg
newData$dayAvg <- NULL

newStepsAday <- with(newData, tapply(steps, date, sum, na.rm = T))
hist(newStepsAday, main = "Number of steps taken each day", xlab = "Number of steps")
```
<br>The mean and median total number of steps taken per day after replacing NA values is: 
```{r}
mean(newStepsAday,   na.rm = T)
median(newStepsAday, na.rm = T)
```
<br>These values do not differ from the estimates from the first part of the assignment because I replaced NAs with "day" averages (If I had replcaed NAs with interval averages, mean and median would have changed slightly). 

### Are there differences in activity patterns between weekdays and weekends?
```{r}
newData["weekend"] <- grepl("S(at|un)", weekdays(as.Date(data$date), abbr = TRUE))
newData[newData$weekend==FALSE,]$weekend <- "weekday"
newData[newData$weekend==TRUE, ]$weekend <- "weekend"

avgs <- aggregate(steps ~ interval + weekend, data = newData, mean)
ggplot(avgs, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(weekend ~ .) +
  xlab("5-minute interval") +
  ylab("number of steps")
```
<br>Yes, there are.